---
title: "Academia to Board Game Brilliance - A Data-Driven Strategy for 'Academia: The Board Game"
author: "Paulo Henrique da Silva Mota"
format: 
    pdf:
      toc: true
      toc-depth: 3
      toc-title: Summary
   
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{graphicx}
  - \usepackage{eso-pic}
  - \usepackage{tikz}
  - \AtBeginDocument{\thispagestyle{empty}\begin{tikzpicture}[remember picture,overlay]
          \node at (current page.south) [yshift=6cm] {\includegraphics[width=0.25\paperwidth,height=0.25\paperheight,keepaspectratio]{logo.png}};
          \node at (current page.center) [yshift=5cm] [anchor=north,yshift=-2cm] {\Huge\textbf{Academia to Board Game Brilliance}};
          \node at (current page.center) [yshift=3cm] [anchor=north,yshift=-2cm] {\huge\textbf{A Data-Driven Strategy for Academia The Board Game}};
          \node at (current page.south) [yshift=5cm] [anchor=north,yshift=-2cm] {\normalsize\textbf{June 2024}};
          \end{tikzpicture}\clearpage}
editor: visual
---

```{r}
#| echo: false
#| warning: false
library(tidytuesdayR)
# Load the required library
library(httr)
library(dplyr)

# Check if the CSV files exist
if (!file.exists("ratings.csv") | !file.exists("details.csv")) {
  # Load the tidytuesdayR library
  library(tidytuesdayR)
  
  # Load the datasets for the specified date
  tuesdata <- tidytuesdayR::tt_load('2022-01-25')
  
  # Assign the datasets to variables
  ratings <- tuesdata$ratings
  details <- tuesdata$details
  
  # Check if the data frames are empty
  if (nrow(ratings) == 0 | nrow(details) == 0) {
    cat("Error: One or both datasets are empty. Cannot export.\n")
  } else {
    # Export the datasets as CSV files
    write.csv(ratings, "ratings.csv", row.names = FALSE)
    write.csv(details, "details.csv", row.names = FALSE)
    
    cat("Datasets have been loaded from tidytuesdayR and exported as 'ratings.csv' and 'details.csv'.\n\n")
  }
} else {
  # Load datasets from CSV files
  ratings <- read.csv("ratings.csv")
  details <- read.csv("details.csv")
  
  cat("Datasets have been loaded from existing CSV files.\n\n")
}
```

# Data Preparation and Description

## Data Source

The data source pertains to BoardGameGeek Reviews and is composed of two datasets. The first dataset contains details about various board games, and the second dataset includes reviews from users for these games, linked by an ID. All reviews are from users, and their comments are included.

```{r}
#| echo: false
#| warning: false

# Get the dimensions of the datasets
ratings_dim <- dim(ratings)
details_dim <- dim(details)

# Print the number of rows and columns for each dataset separately
cat("Ratings Dataset:\n")
cat("Number of rows:", ratings_dim[1], "\n")
cat("Number of columns:", ratings_dim[2], "\n\n")

cat("Details Dataset:\n")
cat("Number of observations:", details_dim[1], "\n")
cat("Number of columns:", details_dim[2], "\n")

```




```{r}
#| echo: false
#| warning: false

# Ratings Table
ratings <- data.frame(
  variable = c("num", "id", "name", "year", "rank", "average", "bayes_average", "users_rated", "url", "thumbnail"),
  class = c("double", "double", "character", "double", "double", "double", "double", "double", "character", "character"),
  description = c("Game number", "Game ID", "Game name", "Game year", "Game rank", "Average rating", "Bayes average rating", "Users rated", "Game url", "Game thumbnail")
)

# Details Table
details <- data.frame(
  variable = c("num", "id", "primary", "description", "yearpublished", "minplayers", "maxplayers", "playingtime", "minplaytime", "maxplaytime", "minage", 
               "boardgamecategory", "boardgamemechanic", "boardgamefamily", "boardgameexpansion", "boardgameimplementation", "boardgamedesigner", 
               "boardgameartist", "boardgamepublisher", "owned", "trading", "wanting", "wishing"),
  class = c("double", "double", "character", "character", "double", "double", "double", "double", "double", "double", "double", 
            "character", "character", "character", "character", "character", "character", "character", "character", "double", "double", "double", "double"),
  description = c("Game number", "Game ID", "Primary name", "Description of game", "Year published", "Min n of players", "Max n of players", "Playing time in minutes", 
                  "Min play time", "Max play time", "minimum age", "Category", "Mechanic", "Board game family", "Expansion", "Implementation", "Designer", 
                  "Artist", "Publisher", "Num owned", "Num trading", "Num wanting", "Num wishing")
)
                

knitr::kable(details,
             format = "html",
             caption = "Board game details dataset")

knitr::kable(ratings,
             format = "html",
             caption = "Ratings dataset")



```


## Data Preparation
Detail the steps taken to clean and preprocess the data in R. Include code snippets where appropriate.

## Performance Measures
Discuss two performance measures (e.g., average rating, number of ratings) and their relevance to the goal of making money.

## Key Variables
Identify and describe three to four key variables that could inform game design actions (e.g., mechanics, play time, number of players).

## Control Variables
Identify variables that might affect game performance but are not part of the strategy options (e.g., year published, game category).

## Data Summary
Provide a table and a figure summarizing the frequency and distributions of relevant variables.

# Data Analysis

## Relationships with Performance Measures
Analyze how key design variables relate to performance measures using appropriate statistical methods.

## Model Evaluation
Discuss the appropriateness of the model based on R output. Explore other methods to test the relationships (e.g., ANOVA, correlation analysis).

## Causality Discussion
Examine the potential for causal inference in the dataset. Discuss limitations and assumptions.

## Results Presentation
Provide a single result table and a figure to present the analysis findings.

# Strategy Recommendation
Strategic Insights
Build on the analysis to provide strategic recommendations for the game design.

## Limitations
Discuss the limitations of the analysis and recommendations.

## Additional Data and Analysis
Suggest additional data that could further inform the strategy and potential analyses that could be conducted with it.

## Conclusion
Summarize the key findings and strategic recommendations. Reinforce the goal of creating a successful and profitable board game.

# References
List all sources and references used in the report.

