---
title: "R Notebook"
output: html_notebook
---

Run -\> *Ctrl+Shift+Enter*.

Add Cell -\> *Ctrl+Alt+I*.

*Preview on HTML -\> Ctrl+Shift+K*

## Reading Data

```{r}
library(tidytuesdayR)

tuesdata <- tidytuesdayR::tt_load('2022-01-25')

ratings <- tuesdata$ratings
details <- tuesdata$details

```

```{r}
head(ratings)

```

Columns Ratings Target : num, id, name, year, rank, average , Bayes_average, users_rated

Columns to drop : url

```{}
```

```{r}
head(details) 
```

Columns to drop : description, boardgamefamily, boardgameexpansion, boardgameimplementation,

Not sure: boardgameartist, boardgamepublisher,

Columns to clean: arrays -\> boardgamescategory, boardgamemecchanic, boardgamedesgin

```{r}
dim(ratings)

```

```{r}
```

```{r}tuition_income %>%}
    dplyr::select(-c(net_cost, income_lvl)) %>%
    dplyr::group_by(name, year) %>%
    dplyr::summarise(avg_per_year = mean(total_price, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(name, year) -> temp_r

```

```{r}
dim(details)

```

```{}
```
