---
title: "R Notebook"
output: html_notebook
---

Run -\> *Ctrl+Shift+Enter*.

Add Cell -\> *Ctrl+Alt+I*.

*Preview on HTML -\> Ctrl+Shift+K*

## Reading Data

```{r}
library(tidytuesdayR)

tuesdata <- tidytuesdayR::tt_load('2022-01-25')

ratings <- tuesdata$ratings
details <- tuesdata$details

# Ensure 'id' columns are integers (or characters)
ratings$id <- as.integer(ratings$id)
details$id <- as.integer(details$id)
# Merge datasets on 'num' and 'id'
merged_data <- merge(ratings, details, by = c("id", "id"))

```

```{r}
head(ratings)

```

Columns Ratings Target : num, id, name, year, rank, average , Bayes_average, users_rated

Columns to drop : url

```{r}
head(details) 
```

Columns to drop : description, boardgamefamily, boardgameexpansion, boardgameimplementation,

Not sure: boardgameartist, boardgamepublisher

Columns to clean: arrays -\> boardgamescategory, boardgamemecchanic, boardgamedesgin

```{r}
dim(ratings)


```

```{r}
colnames(ratings)

```

```{r}
colnames(details)

```

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(stringr)
# Assuming ratings and details datasets are already available in the environment

# Ensure 'id' columns are integers (or characters)
ratings$id <- as.integer(ratings$id)
details$id <- as.integer(details$id)
# Merge datasets on 'num' and 'id'
merged_data <- merge(ratings, details, by = c("id", "id"))

# Plot 1: Distribution of average ratings
plot1 <- ggplot(merged_data, aes(x = average)) +
  geom_histogram(binwidth = 0.1, fill = "darkblue", color = "black") +
  labs(title = "Distribution of Average Ratings",
       x = "Average Rating",
       y = "Frequency") +
  theme_minimal()

# Plot 2: Average rating vs. number of users who rated
plot2 <- ggplot(merged_data, aes(x = users_rated, y = average)) +
  geom_point(alpha = 0.5) +
  labs(title = "Average Rating vs. Number of Users Rated",
       x = "Number of Users Rated",
       y = "Average Rating") +
  theme_minimal() +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) 
  # scale_x_log10()  # log scale for better visualization

# Plot 3: Distribution of games published per year
plot3 <- ggplot(merged_data, aes(x = yearpublished)) +
  geom_histogram(binwidth = 1,fill = "darkblue", color = "black") +
  labs(title = "Distribution of Games Published Per Year",
       x = "Year Published",
       y = "Number of Games") +
  theme_minimal()

# Plot 4: Average rating by category
# Since a game can have multiple categories, we need to handle that
library(tidyr)

category_data <- merged_data %>%
  separate_rows(boardgamecategory, sep = ",") %>%
  mutate(boardgamecategory = str_replace_all(boardgamecategory, "\\[|\\]", "")) %>%
  group_by(boardgamecategory) %>%
  summarise(average_rating = mean(average, na.rm = TRUE),
            count = n()) %>%
  filter(count > 5)  # Filter to show categories with more than 5 games


# Print the plots
print(plot1)
print(plot2)
print(plot3)
```


```{r}

# Number of ratings per board game
num_ratings_per_game <- merged_data %>%
  select(name, users_rated) %>%
  arrange(desc(users_rated))

# Number of ratings per board game with average rating
num_ratings_per_game <- merged_data %>%
  select(name, users_rated, average) %>%
  arrange(desc(users_rated))

# Plot the number of ratings per game (top 20 for readability) with average ratings
plot_ratings_per_game <- ggplot(head(num_ratings_per_game, 20), aes(x = reorder(name, users_rated), y = users_rated)) +
  geom_col(fill = "darkblue") +
  geom_text(aes(label = round(average, 2)), hjust = -0.2, color = "black", size = 3.5) +
  labs(title = "Top 20 Board Games by Number of Ratings",
       x = "Board Game",
       y = "Number of Ratings") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

# Print the plot
print(plot_ratings_per_game)

# Save the plot to a file with increased width
ggsave("plot_ratings_per_game.png", plot = plot_ratings_per_game, width = 12, height = 8)
```

```{r}

# Plot the number of ratings per game (top 20 for readability) with average ratings
plot_ratings_per_game <- ggplot(head(num_ratings_per_game, 20), aes(x = reorder(name, average), y = users_rated)) +
  geom_col(fill = "darkgreen") +
  geom_text(aes(label = round(average, 2)), hjust = -0.2, color = "black", size = 3.5) +
  labs(title = "Top 20 Board Games by Average Rating",
       x = "Board Game",
       y = "Number of Ratings") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

# Print the plot
print(plot_ratings_per_game)

```


```{r}

category_data <- merged_data %>%
  separate_rows(boardgamemechanic, sep = ",") %>%
  mutate(boardgamemechanic = str_replace_all(boardgamemechanic, "\\[|\\]", "")) %>%
  group_by(boardgamemechanic) %>%
  summarise(average_rating = mean(average, na.rm = TRUE),
            count = n()) %>%
  filter(count > 5)  # Filter to show categories with more than 5 games

plot5 <- ggplot(head(category_data,20), aes(x = reorder(boardgamemechanic, average_rating), y = average_rating)) +
  geom_col(fill = "purple") +
  labs(title = "Average Rating by Mechanic",
       x = "Mechanic",
       y = "Average Rating") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

print(plot5)

```
```{r}

category_data <- merged_data %>%
  separate_rows(boardgamefamily, sep = ",") %>%
  mutate(boardgamefamily = str_replace_all(boardgamefamily, "\\[|\\]", "")) %>%
  group_by(boardgamefamily) %>%
  summarise(average_rating = mean(average, na.rm = TRUE),
            count = n()) %>%
  filter(count > 5)  # Filter to show categories with more than 5 games

plot5 <- ggplot(head(category_data,20), aes(x = reorder(boardgamefamily, average_rating), y = average_rating)) +
  geom_col(fill = "purple") +
  labs(title = "Average Rating by Family",
       x = "Family",
       y = "Average Rating") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

print(plot5)


```

```{r}

category_data <- merged_data %>%
  separate_rows(boardgameexpansion, sep = ",") %>%
  mutate(boardgameexpansion = str_replace_all(boardgameexpansion, "\\[|\\]", "")) %>%
  group_by(boardgameexpansion) %>%
  summarise(average_rating = mean(average, na.rm = TRUE),
            count = n()) %>%
  filter(count > 5)  # Filter to show categories with more than 5 games

plot7 <- ggplot(head(category_data,20), aes(x = reorder(boardgameexpansion, average_rating), y = average_rating)) +
  geom_col(fill = "purple") +
  labs(title = "Average Rating by Implementation",
       x = "Implementation",
       y = "Average Rating") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

print(plot7)

```
```{r}

category_data <- merged_data %>%
  separate_rows(boardgamedesigner, sep = ",") %>%
  mutate(boardgamedesigner = str_replace_all(boardgamedesigner, "\\[|\\]", "")) %>%
  group_by(boardgamedesigner) %>%
  summarise(average_rating = mean(average, na.rm = TRUE),
            count = n()) %>%
  filter(count > 5)  # Filter to show categories with more than 5 games

plot7 <- ggplot(head(category_data,20), aes(x = reorder(boardgamedesigner, average_rating), y = average_rating)) +
  geom_col(fill = "purple") +
  labs(title = "Average Rating by Designer",
       x = "Designer",
       y = "Average Rating") +
  theme_minimal() +
  coord_flip()  # Flip coordinates for better readability

print(plot7)

```
```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Select only numeric columns
numeric_data <- merged_data %>% select(where(is.numeric))

# Calculate the correlation matrix
correlation_matrix <- cor(numeric_data)
# Melt the correlation matrix into a long format
melted_corr_matrix <- melt(correlation_matrix)

# Create the correlation heatmap with annotations
heatmap <- ggplot(melted_corr_matrix, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(low = "red", high = "green", mid = "lightyellow",
                         midpoint = 0, limit = c(-1, 1), space = "Lab",
                         name = "Correlation") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(title = "Correlation Heatmap",
         x = "Variables",
         y = "Variables") +
    geom_text(aes(label = round(value, 2)), color = "black", size = 3)

# Display the heatmap
print(heatmap)
```


```{r}

# Load necessary libraries
library(ggplot2)

# Assuming 'merged_data' contains the necessary columns 'average' and 'users_rated'
# Let's plot 'average' rating against 'users_rated'

# Scatter plot
plot <- ggplot(merged_data, aes(x = users_rated, y = average)) +
  geom_point(alpha = 0.5, color = "blue") +  # Add points with transparency
  labs(title = "Average Rating vs. Number of Reviews",
       x = "Number of Reviews (Users Rated)",
       y = "Average Rating") +  # Add labels
  theme_minimal()  # Apply a minimal theme

# Display the plot
print(plot)
```

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)


# Clean and split the categories
clean_data <- merged_data %>%
  mutate(boardgamecategory = gsub("\\[|\\]|'", "", boardgamecategory)) %>%
  separate_rows(boardgamecategory, sep = ", ") %>%
  filter(boardgamecategory != "")

# Count the occurrences of each category
category_distribution <- clean_data %>%
  count(boardgamecategory) %>%
  arrange(desc(n))

# Select the top 20 categories
top20_category_distribution <- head(category_distribution, 30)

# Plot the distribution of all categories
ggplot(category_distribution, aes(x = reorder(boardgamecategory, -n), y = n, fill = n)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8), 
        plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10)) +
  labs(title = "Distribution of Board Game Categories", x = "Category", y = "Frequency") +
  coord_flip()

````